Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-27 paper/ijcai16.tex
paper: modificato esperimenti

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani are
single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a XXXXXX.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. \andrea{@stefano:
  possiamo riportare la mediana anche per i tempi?}  Our {\sc
  SetMargin} algorithm for $k=2$ is reported in orange, Guo in blue,
Viappiani QI in green and Viappiani EUS in gray. For all algorithms,
one iteration corresponds to a single pairwise query to the user.
Results for $r=3$ and $r=4$ are reported.  For dense weight vector
distributions (first two rows), our approach achieves results which
are indistinguishable from the competitors in a fraction of their
time.  Indeed, all Bayesian approaches cannot scale over $r=4$ (for
$r=5$, Guo did not finish a single user after 24 hours for, Viappiani
QI takes 20 minutes per user), \andrea{20 minuti mi paiono pochi, sei
  sicuro?}  while our algorithm can easily scale to much larger
datasets, as will be shown later on. For sparse weight vector
distributions (last two rows) our approach, in addition to being
substantially faster on each iteration, requires less queries in order
to reach optimal solutions. This is an expected result as the
sparsification norm in our formulation ($\| \vw^{i} \|_1$) is
enforcing sparsity in the weight vectors, while none of the other
approaches are explicitly designed to favour sparse functions.

@@

In order to study the effect of increasing the number of weight
vectors in our formulation, we also ran {\sc SetMargin} varying the
parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
results on $r=4$ and $r=5$ datasets for the uniform and normal sparse
distributions\footnote{We picked the most difficult dense distribution
  (uniform) and the simplest sparse distribution (normal) to provide
  the broadest possible picture, as space constraints prevent us from
  including all results. Results for the other distributions are
  however qualitative similar to the ones reported.}, for $k=2$
(orange), $k=3$ (blue) and k=4 (green). The first and third columns
report results in terms of number of iterations. It can be seen that
increasing the number of weight vectors tends to favour earlier
convergence, especially for the more complex dataset ($r=5$). However,
as in each iteration the user is asked to compare $k$ items, different
values of $k$ imply a different cognitive effort for the user. The
second and fourth columns report results in terms of number of
queries, where we count all $k \choose 2$ pairs of queries when
comparing $k$ items. In this case, $k=2$ seems to be the best
option. The cognitive cost for the user will likely lay in between
these two extremes, but formalizing this concept in an efficient query
ordering strategy needs to face the effect of noise.  A modified
sorting algorithm asking only $O(k\log k)$ queries to the user
resulted in a performance worsening, likely because of a cascading
effect of inconsistence feedback. \andrea{@stefano: l'ho rifrasato un
  po', vediamo se ci piace}

@@

A PC configurations is defined by eight attributes: computer type (laptop,
desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)
size, and price.
%See Table~\ref{tab:pcdataset} for an overivew of the attributes.
The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by sum of the price of its components plus an offset due to
branding.
%
Interactions between attributes are expressed as Horn clauses. An example
constraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4''\footnote{Despite the components
included in the dataset, which are a bit outdated, the dataset itself is pretty
realistic.}, which can be encoded
into MILP form as
%
$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
%
Note that mutual exclusivity between all binary variables of the
one-hot encoding of an attribute ensures that only one of
$x_\text{G3}$ and $x_\text{G4}$ can be 1.  The dataset includes
constraints between the following attributes: manufacturer $\to$ type,
manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
size, type $\to$ monitor size, for 16 Horn constraints total. We
do not report the full list due to space limitations.

% 76 features (somma)
% >>> 3*8*37*8*10*10
% 710400

la crossval ogni 5 iterazioni richiede tempo considerevole, che nel
grafico cumulativo del tempo si assomma alle crossval successivie, ampliando
la varianza; in alcuni casi degeneri per alcuni parametri la cosa prende
tempi biblici. in tali casi i risultati sono tipicamente indecente, quindi mettendo
un timeout sui tempi di cross validation e' possibile ridurre questo problema.

\ref{fig:pc}

@@

\stefano{Add figures and comments.}

@@

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        \\
        \hline \hline
        \multicolumn{4}{c}{{\sc Uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Uniform Sparse}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Normal Sparse}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin} and the
    Bayesian competitors. The thick lines represent the median utility loss
    over all users, while the shaded area is the standard deviation. {\sc
    SetMargin} orange, Guo blue, Viappiani QI green, Viappiani EUS gray.}
\end{figure}

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        \multicolumn{2}{c}{{\sc synthetic 4}} &
        \multicolumn{2}{c}{{\sc synthetic 5}}
        \\
        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
        \\
        \hline \hline
        \multicolumn{4}{c}{{\sc uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
        \\
        \hline
        \multicolumn{4}{c}{{\sc normal sparse}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
        \\
        \hline
     \end{tabular}
    }
    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin} with
    $k=2,3,4$, in orange, blue and green respectively.}
\end{figure}

@@

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
        \multicolumn{2}{c}{{\sc loss/time per query}}
        \\
        \multicolumn{4}{c}{{\sc uniform sparse}}
        \\
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc normal sparse}}
        \\
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
    dataset.}
\end{figure}

>>>>

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani (both QI and EUS) are
single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a XXXXXX.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. \andrea{@stefano:
  possiamo riportare la mediana anche per i tempi?}  Our {\sc
  SetMargin} algorithm for $k=2$ is reported in orange, Guo in blue,
Viappiani QI in green and Viappiani EUS in gray. For all algorithms,
one iteration corresponds to a single pairwise query to the user.
Results for $r=3$ and $r=4$ are reported.  For dense weight vector
distributions (first two rows), our approach achieves results which
are indistinguishable from the competitors in a fraction of their
time.  Indeed, all Bayesian approaches cannot scale over $r=4$ (for
$r=5$, Guo did not finish a single user after 24 hours for, Viappiani
QI takes 20 minutes per user), \andrea{20 minuti mi paiono pochi, sei
  sicuro?}  while our algorithm can easily scale to much larger
datasets, as will be shown later on. For sparse weight vector
distributions (last two rows) our approach, in addition to being
substantially faster on each iteration, requires less queries in order
to reach optimal solutions. This is an expected result as the
sparsification norm in our formulation ($\| \vw^{i} \|_1$) is
enforcing sparsity in the weight vectors, while none of the other
approaches are explicitly designed to favour sparse functions.

@@

In order to study the effect of increasing the number of weight
vectors in our formulation, we also ran {\sc SetMargin} varying the
parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
results on $r=4$ and $r=5$ datasets for the uniform and sparse normal
distributions\footnote{We picked the most difficult dense distribution
  (uniform) and the simplest sparse distribution (normal) to provide
  the broadest possible picture, as space constraints prevent us from
  including all results. Results for the other distributions are
  however qualitative similar to the ones reported.}, for $k=2$
(orange), $k=3$ (blue) and k=4 (green). The first and third columns
report results in terms of number of iterations. It can be seen that
increasing the number of weight vectors tends to favour earlier
convergence, especially for the more complex dataset ($r=5$). However,
as in each iteration the user is asked to compare $k$ items, different
values of $k$ imply a different cognitive effort for the user. The
second and fourth columns report results in terms of number of
queries, where we count all $k \choose 2$ pairs of queries when
comparing $k$ items. In this case, $k=2$ seems to be the best
option. The cognitive cost for the user will likely lay in between
these two extremes, but formalizing this concept in an efficient query
ordering strategy needs to face the effect of noise.  A modified
sorting algorithm asking only $O(k\log k)$ queries to the user
resulted in a performance worsening, likely because of a cascading
effect of inconsistence feedback. \andrea{@stefano: l'ho rifrasato un
  po', vediamo se ci piace}

@@

A PC configurations is defined by eight attributes: computer type (laptop,
desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)
size, and price.
%See Table~\ref{tab:pcdataset} for an overivew of the attributes.
The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by sum of the price of its components plus an offset due to
branding.
%
Interactions between attributes are expressed as Horn clauses. An example
constraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4''\footnote{Despite the components
included in the dataset, which are a bit outdated, the dataset itself is pretty
realistic.}, which can be encoded
into MILP form as
%
$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
%
Note that mutual exclusivity between all binary variables of the
one-hot encoding of an attribute ensures that only one of
$x_\text{G3}$ and $x_\text{G4}$ can be 1.  The dataset includes
constraints between the following attributes: manufacturer $\to$ type,
manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
size, type $\to$ monitor size, for 16 Horn constraints total. We do
not report the full list due to space limitations. Note that the
search space is of the order of hundreds of thousands of candidate
configurations, and is far beyond reach of existing Bayesian
approaches. Furthermore, existing Bayesian preference elicitation
approaches are designed for categorical attributes and cannot deal
with (even dependent) continuous features like the price. \andrea{@andrea: mention Bonilla}

Figure~\ref{fig:pc} reports results of {\sc SetMargin} for $k=2$
(red), $k=3$ (blue) and $k=4$ (green) using the sparse uniform (top
row) and sparse normal (bottom row) distributions. The first and third
column report utility loss for increasing number of iterations and
queries respectively, showing a behaviour which is similar to the one
in Figure~\ref{fig:selfcomparison}. Overall, between 50 and 70 queries
on average are needed in order to find out a solution which is only
10\% worse than the optimal one, out of the more than 700,000
thousands available. \andrea{something about the lack of smart
  constraints, as a vendor would have..}  The second and fourth
columns report cumulative times. Note that in some cases, standard
deviations have a bump. This is due to cases in which some of the
hyperparameters of the internal cross validation result in
ill-conditioned optimization problems which are hard to solve. These
exceptions can be easily dealt with by setting an appropriate timeout
on the cross validation without affecting the results, as these
hyperparameters typically end up having bad performance and being
discarded.

@@

@@

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        \\
        \hline \hline
        \multicolumn{4}{c}{{\sc Uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Sparse Uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Sparse Normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
      and the Bayesian competitors on the $r=3$ (left) and $r=4$
      (right) datasets. Each row represents a different distribution
      for sampling the underlying user utility.  The x-axis reports
      the number of iterations (i.e. pairwise queries), while the
      y-axis reports the utility loss (first and third columns) and
      the cumulative time in seconds (second and fourth
      columns). Thick lines indicate median values over all users,
      while standard deviations are shown as shaded areas. Curves
      represent {\sc SetMargin} (orange), Guo (blue), Viappiani QI
      (green) and Viappiani EUS (gray) respectively. \andrea{@stefano remove median and average from y-axis}}
\end{figure}

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        \multicolumn{2}{c}{{\sc synthetic 4}} &
        \multicolumn{2}{c}{{\sc synthetic 5}}
        \\
        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
        \\
        \hline \hline
        \multicolumn{4}{c}{{\sc uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
        \\
        \hline
        \multicolumn{4}{c}{{\sc sparse normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
        \\
        \hline
     \end{tabular}
    }
    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
      with $k=2,3,4$, in orange, blue and green respectively for the
      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions for user utility. Median (thick lines) and standard deviation (shaded areas) utility loss values are reported for increasing number of iterations (first and third columns) and pairwise queries (second and fourth columns).}
\end{figure}

@@

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
        \multicolumn{2}{c}{{\sc loss/time per query}}
        \\
        \multicolumn{4}{c}{{\sc sparse uniform}}
        \\
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc sparse normal}}
        \\
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
    dataset.}
\end{figure}
